---
title: "Homework 3"
author: "Al Freeman"
date: "2/23/2022"
output: html_document
---
Question 1: I found spaces that were greater than two to keep spacing between 'First String' and "More Text' uisng /s{2}. I replaced this with a comma 

```{r}
#Find: \s{2,}
#replace: ,
```

Question 2: I found and captured the first word, second word, and "all the rest". I replaced in order of the second capture, first capture, and third capture but I added parentheses around the third capture using a backslash before the third capture call

```{r}
#Find: (\w+), (\w+), (.*)
#Replace: \2 \1 \(\3\)

```

Question 3: I captured .mp3 and found the space following it. I then kept the first capture and replaced the space with a line break

```{r}
#Find: (.mp3)\s
#Replace: \1\r
```

Question 4: I broke the line down piece by piece and captured the first set of integers, then the song title, and separately captured .mp3. I then reordered the captures so that they would reflect the desired outcome 

```{r}
#Find: (\d+)\s(.+)(.mp3)
#Replace: \2_\1\3

```

Question 5: I captured the first letter using \w and then highlighted the rest of the word and comma so that these could be replaced with an underscore. I then captured the species name and isolated the first number uisng /d+./d, so that this would get replaced. I captured the last number so that it would follow the underscore 
```{r}
#Find: (\w)\w+,(\w+),\d+.\d,(\d+)
#Replace: \1\_\2,\3
```

Question 6: The only change I made from the above code was specifying to capture just the first 4 letters of the species name using (\w{1,4})

```{r}
#Find: (\w)\w+,(\w{1,4})\w+,\d+.\d,(\d+)
#Replace: \1\_\2,\3

```

Question 7: I captured the first three of the word using (\w{1,3}) and then highlighted the rest of the word with \w+, repeated this for the species name, and then broke the date down into parts so that the month and day were captured together and the year was captured separately. I then reordered my captures and added commas and spaces where appropriate 

```{r}
#Find: (\w{1,3})\w+,(\w{1,3})\w+,(\d+.\d),(\d+)
#Replace: \1\2, \4, \3

```