---
title: "Homework_10"
author: "Al Freeman"
date: '2022-04-06'
output: html_document
---

For loops and randomization tests
Using a forloop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the matrix. Finally, use return(counter) for the output.

```{r}

library(tidyverse)
# create a numeric vector containing some 0's
my_vec <- rpois(100, lambda = 2)

# -------------------------------
# FUNCTION count_zeros
# description: counts number of zeros in numeric vector
# inputs: numeric vector
# outputs: number of zeros in vector
###################################
count_zeros <- function(vec = rnorm(n = 10, mean = 0, sd = 1)){

# counter variable 0
counter <- 0

  for(i in seq_along(vec)){
  if(vec[i] == 0){
    counter <- counter + 1
  }

}

  return(counter)
} # end of count_zeros
#--------------------------------

# test 1
count_zeros()
## [1] 0
# test 2
count_zeros(vec = my_vec)
## [1] 12

```

Use subsetting instead of a loop to rewrite the function as a single line of code.

```{r}
length(my_vec[my_vec == 0])

```
Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.


```{r}
# -------------------------------
# FUNCTION matrix_fun
# description: creates a matrix of certain dimensions where each element is the product of row * column number
# inputs: two integers (row and column)
# outputs: matrix
###################################
matrix_fun <- function(x  = 2, y = 3){

mat1 <- matrix(nrow = x, ncol = y)

for(i in 1:nrow(mat1)){
  for(j in 1:ncol(mat1)){
    mat1[i,j] <- i * j
  }
}

return(mat1)

} # end of matrix_fun
#--------------------------------

# test function
matrix_fun()
##      [,1] [,2] [,3]
## [1,]    1    2    3
## [2,]    2    4    6
# test with different values
matrix_fun(x = 3, y = 5)


```
In the next few lectures, you will learn how to do a randomization test on your data. We will complete some of the steps today to practice calling custom functions within a for loop. Use the code from the March 31st lecture (Randomization Tests) to complete the following steps:

Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.



```{r}
group <- as.factor(c(rep("A", 30), rep("B", 30), rep("C", 30))) #make group vector
A <- rnorm(n = 30, mean = 22, sd= 2.5) #make corresponding data for group A
B <- rnorm(n = 30, mean = 180, sd = 3)#make corresponding data for group B
C <- rnorm(n = 30, mean = 100, sd = 1)#make corresponding data for group C
response <- c(A, B, C) #combine data into one vector for the responses
newdf <- data.frame(group, response) # combine response vector and groups into a data frame
head(newdf)
```

Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.

```{r}
reshuffle <- function(df = newdf) {
  colnames(df) <- c("group", "response")
  df$response <- sample(df$response, length(df$response), replace = FALSE)

    df2 <- df %>%
     group_by(group) %>%
     summarize(mean = mean(response))
    return(as.vector(df2$mean))
}
reshuffle()


```

Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.


```{r}
#create empty vectors to fill in for loop
rep <- rep(NULL, 100)
grp_meanA <- rep(NULL, 100)
grp_meanB <- rep(NULL, 100)
grp_meanC <- rep(NULL, 100)
df <- data.frame(rep, grp_meanA, grp_meanB, grp_meanC)

#For loop to reshuffle data 100 times and save output as a dataframe-print dataframe to chekc there are 4 columns 
for (i in 1:100) {
  df[i,1] <- i
  for (j in 1:3) {
  df_reshuff <- data.frame(c(A, B, C), reshuffle())
  df[i,j+1] <- df_reshuff[j,2]
  
  } # end j for loop
} #end i for loop
colnames(df) <- c("rep", "grp_meanA", "grp_meanB", "grp_meanC")
head(df)#four columns 

```

Use qplot() to create a histogram of the means for each reshuffled group. Or, if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?

```{r}
df_all <- df %>%
  pivot_longer(cols = grp_meanA:grp_meanC, names_to = "group", values_to = "Values")

ggplot(df_all, aes(x=Values)) + geom_histogram() + facet_wrap(~group)

#the original had higher counts for the means


```